---
title: "FGV-EEPGE Macro presentation plots"
output: html_notebook
---

```{r}
rm(list=ls())
library('dplyr')
library('ggplot2')
library("readxl")
library("lubridate")
library('padr')
library('here')

source(file.path(getwd(), 'plots', 'plot_funcs.R'))
source(file.path(getwd(), 'models', 'utils.R'))
source("https://gist.github.com/ellisp/4002241def4e2b360189e58c3f461b4a/raw/0e5dc52262e86b5d21b914db3553e58b0b8d1d20/dualplot.R")

INPUT_PATH <- file.path(here(), 'src', 'data', 'inputs')
OUTPUT_PATH <- file.path(here(), 'src', 'data', 'outputs')
UTILS_PATH <- file.path(here(), 'src', 'data', 'utils')
PPT_OUTPUT_PATH <- file.path(OUTPUT_PATH, "eepge-macro-presentation")
```

# Load MAS interventions, exchange rates, trade balance data and DLM betas datasets

```{r}
# mas_decisions_df <- read_excel(file.path(UTILS_PATH, "mas_interventions.xlsx"))
currencies_df <- read.csv(file.path(INPUT_PATH, "daily-currencies.csv"))
trade_df <- read.csv(file.path(INPUT_PATH, "relative-trade-data.csv"))

model <- 'rolling-ols'
scale_type <- 'rolling_scale'
output_reference <- paste0(scale_type)
freq <- "weekly"
window_size <- 48
path_rolling_dlm <- file.path(OUTPUT_PATH, 
                              model,
                              output_reference,
                              paste0("model_results_", 
                              freq, "_", scale_type, "_", window_size, ".rds"))
model_output <- readRDS(path_rolling_dlm)

if (model == "dlm"){
  betas_df <- model_output$smooth$s
}else if (model == "rolling-ols"){
  betas_df <- model_output$model$coefs
}
```

# Make the proper adjustments on column names, date format etc

```{r}
# RENAME
name_list <- list(KRW="Korea",
                  MYR="Malaysia",
                  CNY="China",
                  THB="Thailand",
                  IDR="Indonesia",
                  TWD="Taiwan",
                  INR="India",
                  JPY="Japan",
                  EUR="EU",
                  AUD="Australia",
                  GBP="England",
                  SGD="Singapore")
# mas_decisions_df <- mas_decisions_df %>% mutate(date=ymd(date))
currencies_df <- currencies_df %>% mutate(date=ymd(date))
for (colname in names(name_list)){
  new_name <- name_list[[colname]]
  
  # currencies
  orig_name <- colnames(currencies_df)[grep(paste0("^", colname), colnames(currencies_df))]
  currencies_df <- currencies_df %>% rename(!!new_name := !!orig_name)
  
  # betas of the dlm model
  orig_name <- colnames(betas_df)[grep(paste0("^", colname), colnames(betas_df))]
  betas_df <- betas_df %>% rename(!!new_name := !!orig_name)
}
trade_df <- trade_df %>% mutate(date=ymd(date))

# RESAMPLE
currencies_df <- resample_data(currencies_df %>% select(-USD.Curncy) %>% filter(date >= "2006-01-01"))
```

# 1) Beta Estimates
## DLM estimate vs trade proxy

```{r}
apply_mean <- TRUE
apply_zscore <- TRUE
start_date <- "2008-01-01"

for (colname in colnames(betas_df %>% select(-date, -intercept))){
  
  # select data and merge mixe frequencies
  tmp_betas_df <- betas_df %>% as.data.table() %>% select(date, sym(colname)) %>%
    rename(!!paste0(colname, " beta") := !!paste0(colname))
  tmp_trade_df <- trade_df %>% as.data.table() %>% select(date, sym(colname)) %>%
    rename(!!paste0(colname, " trade") := !!paste0(colname))
  setkey(tmp_betas_df, date)
  
  if (apply_mean == TRUE){
    tmp_trade_df <- tmp_trade_df %>%
      mutate(!!colnames(tmp_trade_df)[2] := roll_mean(!!sym(colnames(tmp_trade_df)[2]), width = 6))
  }
  
  tmp_plot_df <- tmp_betas_df[tmp_trade_df, roll = "nearest" ] %>% drop_na()
  tmp_plot_date <- tmp_plot_df$date
  
  if (apply_zscore == TRUE){
    tmp_plot_df <- lapply(tmp_plot_df, scale) %>% as.data.table()
    tmp_plot_df$date <- tmp_plot_date
    tmp_plot_df <- tmp_plot_df %>% select(date, everything()) 
    tag <- "_zscore"
  }else{
    tag <- ""
  }
  
  # multiaxis plot data
  plot_img <- ggplot(tmp_plot_df %>% filter(date >= start_date), aes(x=date)) + 
  geom_line(aes(y=!!sym(colnames(tmp_plot_df)[2]), colour="Beta estimates")) + 
  geom_line(aes(y=!!sym(colnames(tmp_plot_df)[3]), colour="Trade proxy")) + 
  scale_y_continuous(name = "", sec.axis = sec_axis(~.*1, name="")) + 
  ggtitle(colname)

  # save plot
  ggsave(plot = plot_img,
       filename = file.path(PPT_OUTPUT_PATH, paste0(colname, "_dlm_betas_trade", tag,".png")),
       width = 12,
       height = 8,
       dpi = 100)
}
```

## All DLM estimates

```{r}
plot_img <- ggplot(data = betas_df %>% filter(date >= start_date) %>% melt("date"), 
                   mapping = aes(x = date, y = value, colour = variable, group = variable)) +
              geom_line() +
              scale_x_date(date_breaks = "1 year") + 
              theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
              ggtitle("DLM parameter estimates (with intercept")

# save plot
ggsave(plot = plot_img,
     filename = file.path(PPT_OUTPUT_PATH, "all_betas_dlm_estimates_intercept.png"),
     width = 12,
     height = 8,
     dpi = 100)
plot_img
```

```{r}
plot_img <- ggplot(data = betas_df %>% select(-intercept) %>% filter(date >= start_date) %>% melt("date"), 
                   mapping = aes(x = date, y = value, colour = variable, group = variable)) +
              geom_line() +
              scale_x_date(date_breaks = "1 year") + 
              theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
              ggtitle("DLM parameter estimates (without intercept)")

# save plot
ggsave(plot = plot_img,
     filename = file.path(PPT_OUTPUT_PATH, "all_betas_dlm_estimates.png"),
     width = 12,
     height = 8,
     dpi = 100)
plot_img
```

# 2) Backtest

## EWMA vol bands

```{r}
strategy <- "ewma"

ewma_outputs <- readRDS(file.path(OUTPUT_PATH,
                                  model,
                                  output_reference,
                                  paste0("backtest_", freq, "_", scale_type, "_", window_size, "_", strategy, "_results.rds")))
ewma_rets_df <- ewma_outputs$returns %>% drop_na()
ewma_signal_df <- ewma_outputs$signal
ewma_position_df <- ewma_outputs$positions
ewma_bars_ret_df <- ewma_outputs$bars_ret
ewma_bars_df <- ewma_outputs$bars

ewma_cumret_df <- cumprod(1+(ewma_rets_df %>% select(-date))) %>% mutate(date=ewma_rets_df$date) %>% select(date, everything())
```

### Cummulative returns of each currency

```{r}
plot_img <- ggplot(data = ewma_cumret_df %>% filter(date >= start_date) %>% melt("date"), 
                   mapping = aes(x = date, y = value, colour = variable, group = variable)) +
              geom_line() +
              scale_x_date(date_breaks = "1 year") + 
              theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
              ggtitle("Cummulative returns of $1 invested in each asset")

# save plot
ggsave(plot = plot_img,
     filename = file.path(PPT_OUTPUT_PATH, paste0("cumret_all_assets_", model, "_", strategy, ".png")),
     width = 12,
     height = 8,
     dpi = 100)
plot_img
```

### Cummulative returns of the portfolio

```{r}
plot_img <- ggplot(data = data.frame(date=ewma_rets_df$date,
                                     rets=cumprod(1+rowSums(ewma_rets_df %>% select(-date)))),
                   mapping = aes(x = date, y = rets)) + 
              geom_line() + 
              scale_x_date(date_breaks = "1 year") + 
              theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
              ggtitle("Cummulative returns of $1 invested in the portfolio")

# save plot
ggsave(plot = plot_img,
     filename = file.path(PPT_OUTPUT_PATH, paste0("cumret_portfolio_", model, "_", strategy, ".png")),
     width = 12,
     height = 8,
     dpi = 100)
plot_img
```

### Cummulative returns of the portfolio with vol. target of 5%

```{r}
ewma_ret_df <- data.frame(date=ewma_rets_df$date,
                          portfolio=rowSums(ewma_rets_df %>% select(-date)))
roll_portfolio_vol <- roll_sd(x = ewma_ret_df$portfolio, width = 12) * sqrt(52) * 100
roll_portfolio_vol <- 5 / roll_portfolio_vol
vol_adj_ret_df <- ewma_ret_df %>% mutate(portfolio=portfolio*roll_portfolio_vol) %>% drop_na()

plot_img <- ggplot(data = data.frame(date=vol_adj_ret_df$date,
                                     portfolio=cumprod(1+vol_adj_ret_df %>% select(-date))),
                   mapping = aes(x = date, y = portfolio)) + 
              geom_line() + 
              scale_x_date(date_breaks = "1 year") + 
              theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
              ggtitle("Cummulative returns of $1 invested in the portfolio (vol. target 5%)")

plot_img2 <- ggplot(data = data.frame(date=vol_adj_ret_df$date,
                                      vol=roll_sd(x = vol_adj_ret_df$portfolio, width = 12) * sqrt(52) * 100),
                    mapping = aes(x = date, y = vol)) +
              geom_line() + 
              scale_x_date(date_breaks = "1 year") + 
              theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
              ggtitle("Realized volatility of the portfolio")

print(paste0("Sharpe ratio of ", as.character(round(mean(vol_adj_ret_df$portfolio, na.rm = TRUE) / sd(vol_adj_ret_df$portfolio, na.rm = TRUE) * sqrt(52), 2))))

# save plot
ggsave(plot = plot_img,
     filename = file.path(PPT_OUTPUT_PATH, paste0("cumret_portfolio_vol_target_", model, "_", strategy, ".png")),
     width = 12,
     height = 8,
     dpi = 100)
plot_img
plot_img2
```

### Positions scaling factor

```{r}
mean(roll_portfolio_vol, na.rm = TRUE)
ggplot(data = data.frame(date=ewma_ret_df$date,
                         scaling_factor=roll_portfolio_vol) %>% drop_na(),
                   mapping = aes(x = date, y = scaling_factor)) + 
              scale_x_date(date_breaks = "1 year") + 
              theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
              geom_line() 
```

### Sanity check assets ann. mean and vol

```{r}
(apply(ewma_bars_ret_df %>% select(-date), 2, sd) * sqrt(252))
```

### Portfolio statistics

```{r}
ewma_portfolio_rets_df <- rowSums(ewma_rets_df %>% select(-date)) %>% as.data.table() %>%
  mutate(date=ewma_rets_df$date) %>% select(date, everything()) %>% rename(portfolio=".")
ewma_portfolio_rets_ts <- as.xts(ewma_portfolio_rets_df$portfolio, order.by = ewma_portfolio_rets_df$date)
ewma_stats <- list(mean=mean(ewma_portfolio_rets_df$portfolio) * 52,
                   vol=sd(ewma_portfolio_rets_df$portfolio) * sqrt(52),
                   SR=(mean(ewma_portfolio_rets_df$portfolio) / sd(ewma_portfolio_rets_df$portfolio)) * sqrt(52))
as.data.table(ewma_stats)
```

### Intuition why the strategy works

```{r}
tmp_ewma_position_df <- ewma_position_df %>% select(date, SGD) %>%
  filter((date >= "2015-01-01")&(date <= "2016-01-01")) %>% rename(positions=SGD)
tmp_ewma_bars_df <- ewma_bars_df %>% select(date, SGD) %>%
  filter((date >= "2015-01-01")&(date <= "2016-01-01")) %>% rename(SGD_price=SGD)
tmp_ewma_bars_ret_df <- ewma_bars_ret_df %>% select(date, SGD) %>%
  filter((date >= "2015-01-01")&(date <= "2016-01-01")) %>% rename(SGD_ret=SGD)
tmp_ewma_signal_df <- ewma_signal_df %>% filter((date >= "2015-01-01")&(date <= "2016-01-01"))
tmp_ewma_rets_df <- ewma_rets_df %>% select(date, SGD) %>%
  filter((date >= "2015-01-01")&(date <= "2016-01-01")) %>% rename(portfolio_ret=SGD)

tmp_check_strategy <- merge(tmp_ewma_signal_df, tmp_ewma_position_df, by = "date") %>% merge(tmp_ewma_bars_df, by = "date")  %>% merge(tmp_ewma_bars_ret_df, by = "date") %>%  merge(tmp_ewma_rets_df, by = "date")
tail(round(x = tmp_check_strategy %>% select(-date), digits = 3) %>%
       mutate(date=tmp_check_strategy$date) %>% select(date, everything()))
```

























